---
title: "Decision Trees & Random Forest"
author: "Niti Wattanasirichaigoon (A20406934)"
date: "12/4/2020"
output: pdf_document
---


# Load Data
```{r}
df_train <- read.csv(file="H1.csv",sep=",",stringsAsFactors = FALSE, na.strings = "NULL")
df_test <- read.csv(file="H2.csv",sep=",",stringsAsFactors = FALSE, na.strings = "NULL")
```


# Data Transformation for training and testing
```{r}
df_train$Country <- as.numeric(factor(df_train$Country))
df_train$MarketSegment <- as.numeric(factor(df_train$MarketSegment))
df_train$DepositType <- as.numeric(factor(df_train$DepositType))
df_train$CustomerType <- as.numeric(factor(df_train$CustomerType))
df_train$IsCanceled <- factor(df_train$IsCanceled)

df_train <- df_train[c('IsCanceled','LeadTime', 'Country', 'MarketSegment', 
                       'DepositType', 'CustomerType', 'RequiredCarParkingSpaces',
                       'ArrivalDateWeekNumber')]
                     
df_test$Country <- as.numeric(factor(df_test$Country))
df_test$MarketSegment <- as.numeric(factor(df_test$MarketSegment))
df_test$DepositType <- as.numeric(factor(df_test$DepositType))
df_test$CustomerType <- as.numeric(factor(df_test$CustomerType))
df_test$IsCanceled <- factor(df_test$IsCanceled)

df_test <- df_test[c('IsCanceled','LeadTime', 'Country', 'MarketSegment', 
                       'DepositType', 'CustomerType', 'RequiredCarParkingSpaces',
                       'ArrivalDateWeekNumber')]

addmargins(table(df_train$IsCanceled))
```
# Training and validation split
```{r}
library(caret)
set.seed(111)
trainIndex = createDataPartition(df_train$IsCanceled, p = .8, list = FALSE)
val_x = df_train[-trainIndex, -1]
val_y = df_train[-trainIndex, 1]
```

# Decision Tree
Build simple decision tree, plot it, and calculate misclassification rate.
```{r}
library(rpart)
library(rpart.plot)

# using all attributes
fit = rpart(IsCanceled ~ ., data = df_train[trainIndex,])
rpart.plot(fit, type=3)
```
## Training misclassification rate and confusion matrix
```{r}
tree_pred = predict(fit, df_train[trainIndex,-1], type = 'class')
miss_rate = sum(tree_pred != df_train[trainIndex,1])/length(tree_pred)
miss_rate
```

```{r}
library(e1071)
confusionMatrix(tree_pred, df_train[trainIndex,1])
```


## Validation misclassification rate and confusion matrix
```{r}
tree_pred = predict(fit, val_x, type = 'class')
miss_rate = sum(tree_pred != val_y)/length(tree_pred)
miss_rate
```
```{r}
confusionMatrix(tree_pred, val_y)
```

## Test set misclassification rate and confusion matrix
```{r}
tree_pred = predict(fit, df_test[,-1], type = 'class')
miss_rate = sum(tree_pred != df_test[,1])/length(tree_pred)
miss_rate
```

```{r}
confusionMatrix(tree_pred, df_test[,1])
```


# Random Forest

```{r message=FALSE}
library(randomForest)
set.seed(123)
rf_fit = randomForest(IsCanceled ~ ., data = df_train[trainIndex,], na.action=na.omit)
```
## Training confusion matrix
```{r}
rf_pred = predict(rf_fit, df_train[trainIndex,-1], type = 'class')
confusionMatrix(rf_pred, df_train[trainIndex,1])
```
## Validation confusion matrix
```{r}
rf_pred = predict(rf_fit, val_x)
confusionMatrix(rf_pred, val_y)
```

## Test set misclassification rate and confusion matrix
```{r}
# make predictions
rf_pred = predict(rf_fit, df_test[,-1], type = 'class')
confusionMatrix(rf_pred, df_test[,1])
```


